troubleshooting

### 2024-04-22 README.Rmd
added function to calculate all possible cleavages.

## Workflow

So how does the msps data normalization process work?

1. Takes two input files from PEAKS and combines them
3. Normalizes values and then does a reverse log2 transformation
4. Looks for outliers across replicates. Removes them. 
5. Imputes data for missing values (likely to be very low intesity).
6. Figures out the locations of the detected clevages within the library of peptide sequences used. Is it cleaved at the N or C terminus, or both?

This package also contains tools for doing some basic statistics and data visualizations.

7. Calculates the fold change and the p/q value across experimental conditions using T-tests.   
8. Conducts ANOVA tests to determine if there are any peptides that are significantly different across all time points.  
9. Conducts a hierarchical clustering analysis to visualize overall patterns in the data.

@@

### Loading the files ###
lfq_filename = "test/protein-peptides-lfq.csv"
#file "protein-peptides.csv" exported from PEAKS identification
id_filename = "test/protein-peptides-id.csv"

@@

Before we normalyze data, we need to know what samples are in what groups. We can do that by defining a design matrix. 
```{r}
design_matrix = readr::read_csv("test/design_matrix.csv")

@@

```{r}
design_matrix = readr::read_csv("test/design_matrix.csv")
outliers = handle_outliers(normalyzed_data,design_matrix)

@@

### polishing

@@

### Conducting stats

@@

#### T tests

@@

#### log2FC

@@

#### log2fc_t_tests. 
Sometimes (such as when you want to make volcano plots) it is useful to look at the log2fc and the t test statistics at the same time. This function does that.

@@

#### ANOVA

@@

### Common Data Visualizations

@@

#### PCA

@@

mspms::plot_pca(prepared_for_stats)
```

#### Hierchical clustering

@@

#### Ploting time course.

>>>>

## Overview

There are 4 different types of functions in this package.  Those that are involved in:

1. Making mspms generically useful. These functions are focused on making mspms more generally useful to a wider audience outside of the very specific workflow traditionally used in the O’Donoghue lab. Allows for the use of other types of upstream proteomic data processing, different peptide libraries, etc.

2. Data processing/ normalization. These functions allow the user to normalize and process the MSP-MS data.

3. Statistics. These methods allow the user to perform basic statistics on the normalyzed/processed data.

4. Data visualization. These functions allow the user to visualize the data in a number of useful way.

**Making mspms generically useful**.

1. *calculate_all_cleavages()*: Calculates all possible cleavages for peptide library sequences.

**Data Proessing/ Normalization**.

1. *prepare_for_normalyzer()*:Takes two input files from PEAKS and combines them.    
2. *normalyze()*: Normalizes values and then does a reverse log2 transformation.    
3. *handle_outliers()*: Looks for outliers across replicates. Removes them.     
4. *impute()*: Imputes data for missing values (not including NAs introduced by handle_outliers()).    
5. *join_with_library()*: Joins the normalyzed data with the peptide library sequences.  
6. *add_cleavages()*: Figures out the locations of the detected clevages within the library of peptide sequences used. Is it cleaved at the N or C terminus, or both?  Uses *cterm_cleavage()* and *nterm_cleavage()* to do this.  
7. *polish()*: combines the cleavage info into one column. Discards peptides that were cleaved on both sides or not at all.  
8. *prepare_for_stats()*: Reshapes the data into a long format and appends the data in the design matrix to make statistical testing easy to perform.

**Statistics**.   
1. *log2fc_t_test()*: Calculates the fold change and the p/q value across experimental conditions using T-tests.      
2. *mspms_anova()*: Conducts ANOVA tests to determine if there are any peptides that are significantly different across all time points.

**Data Visualization**.   
1. *plot_heatmap()*: Conducts hierarchical clustering analysis and plots an interactive heatmap to visualize overall patterns in the data.    
2. *plot_pca()*: PCA analysis to visualize the data in a 2D space.  
3. *plot_time_course()*: Plot peptides over time by condition.  
4. *plot_cleavage_motif()*: Visualize the sequence specificity of the cleavage sites. Designed to be similar to what is implemented in IceLogo.

## Making Generically Usefull

### Calculating all cleavages

We might want to calculate all possible cleavages for the peptide library sequences. This is useful for downstream analysis, especially when we are looking at the specificity of the cleavage sites via  as this requires a background of all possible cleavages

We can do this by specifying the number of amino acids after the cleavage site that we are interested in. First lets try 4, which is the default.

```{r}
all_peptide_sequences = mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
                        n_AA_after_cleavage=4)
head(all_peptide_sequences)

We could also try 5 AA after each cleavage site

```{r}
all_peptide_sequences = mspms::calculate_all_cleavages(mspms::peptide_library$library_real_sequence,
                        n_AA_after_cleavage=5)
head(all_peptide_sequences)
```

## Data Normalization/ Processing

@@

### Loading the files ###
lfq_filename = "tests/protein-peptides-lfq.csv"
#file "protein-peptides.csv" exported from PEAKS identification
id_filename = "tests/protein-peptides-id.csv"

@@

Before we normalyze data, we need to know what samples are in what groups. We can do that by defining a design matrix. 
```{r}
design_matrix = readr::read_csv("tests/design_matrix.csv")

@@

```{r}
design_matrix = readr::read_csv("tests/design_matrix.csv")
outliers = handle_outliers(normalyzed_data,design_matrix)

@@

### Polishing

@@

## Statistics

@@

### T tests

@@

### log2FC

@@

### log2fc_t_tests.

Sometimes (such as when you want to make volcano plots) it is useful to look at the log2fc and the t test statistics at the same time. This function does that.

@@

### ANOVA

@@

## Common Data Visualizations

@@

### Plotting sequence specificity motif

Here use an approach similar to what is implemented in ICELogo to visualize the sequence specificity of the cleavage sites. This was used as reference to build the code: https://iomics.ugent.be/icelogoserver/resources/manual.pdf.

```{r}
cleavage_seqs = mspms::prepared_for_stats %>%
  mspms::polish() %>% 
  filter(condition == "DMSO",time == 240) %>%
  pull(cleavage_seq)

background_universe = mspms::all_possible_8mers_from_228_library

mspms::plot_cleavage_motif(cleavage_seqs,background_universe)
```

```{r}
test_seq = readLines("tests/cleavage_seq_test.txt")

mspms::plot_cleavage_motif(test_seq,background_universe)
```

### PCA

@@

```{r}
mspms::plot_pca(prepared_for_stats)
```

### Hierchical clustering

@@

### Ploting time course.
